---
title: "Taking A Look at RCMIP II Hector Results"
author: "Kalyn Dorheim"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE, message = FALSE}
BASE_DIR <- here::here()
OUT_DIR  <- file.path(BASE_DIR, 'output', 'rcmipII', 'netcdfs')
knitr::opts_chunk$set(echo = TRUE,  fig.width = 8, fig.height = 5)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ncdf4)
library(data.table)

nc_files <- list.files(OUT_DIR, 'convert', full.names = TRUE)
```


Let's take a look at what the netcdf file looks like. 

```{r}
nc <- nc_open(nc_files[[1]])
nc 
nc_close(nc)
```

Define a function that will extract and format the different data sets in prepration for plotting results. 

```{r}
lapply(nc_files, function(f){
  
  nc <- nc_open(f)
  
  ensemble <- as.character(0:9999)
  time     <- 1750:2100 # we know that this is going to be the time but we will only want to keep a subset of them
  yr_to_keep <- floor(seq(from=min(time), to=max(time), length.out = 50)) 
  
  vars <- c('Surface Air Temperature Change', 'Surface Air Ocean Blended Temperature Change', 
            'Effective Radiative Forcing|Anthropogenic|Aerosols')
  
  lapply(vars, function(v){
    
    var_output <- as.data.table(ncvar_get(nc, v))
    names(var_output) <- ensemble
    var_output[['time']] <- time
    var_output <- var_output[time %in% yr_to_keep, ]
    
    long <- melt(var_output, id.vars = 'time', variable.name = 'ensemble', value.name = 'value')
    long[['variable']] <- v
    
    return(long)
    
  }) %>%  
    dplyr::bind_rows()-> 
    data
  
  data$scenario <- ncatt_get(nc, 0)$scenario
  
  return(data)
  
  
}) %>% 
  dplyr::bind_rows() ->
  data
```


```{r}
to_plot <- data[ , .("min" = min(value), "max" = max(value), "mean" = mean(value)), by = c("variable", "scenario", "time")]
```

## Plots

The  ribbon is the min/max hector results the dark black line is the Hector ensemble mean. 

```{r}
vars <- unique(to_plot$variable)
plot_list <- list()
for(v in vars){
  
  ggplot(data = to_plot[variable == v, ] ) + 
    geom_ribbon(aes(x = time, ymin = min, ymax = max),alpha = 0.5) +
    geom_line(aes(x = time, y=mean),size =2) +
    facet_wrap('scenario', scales= 'free') + 
    labs(title = paste0('RCMIP  II: ', v)) -> 
    plot
  
  plot_list[[v]] <- plot
  
}
```


```{r}
plot_list$`Surface Air Temperature Change`
```


```{r}
plot_list$`Surface Air Ocean Blended Temperature Change`
```


Let's compare the two different types of temperature. 


```{r, fig.width = 8, fig.height = 5}
 ggplot(data = to_plot[variable %in% c('Surface Air Temperature Change',
                                       'Surface Air Ocean Blended Temperature Change' ), ] ) + 
    geom_ribbon(aes(x = time, ymin = min, ymax = max, fill = variable, color = variable),alpha = 0.5) +
    geom_line(aes(x = time, y=mean, color = variable)) +
    facet_wrap('scenario', scales = 'free') + 
    labs(title = paste0('RCMIP  II: ', v)) -> 
    plot
  plot 
```


As we expected the mean surface temperature is cooler than the mean air temperature, and this difference is more pronounced in the future/higher $CO_2$ scenarios. 


```{r}
plot_list$`Effective Radiative Forcing|Anthropogenic|Aerosols`
```

It makes snese that there is no uncertainty in the 1pctCO2, abrupt2xCO2 and the abrupt4xCO2 runs because they aren't actually driven with aerosols. 
